---
title: "NHANES Research Analysis"
output: html_document
date: "2025-04-17"
---
Race
Household Income
Poverty Ratio
Supermarket Spending 
Disability
Food Security

Hypothesis: There is a correlation between the race a disabled person identifies as and their economic quality of life. 
```{r}
NHANESResearch <- read.csv("/Users/oliviaslaptop/Documents/Data/clean_data")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(modelsummary)
library(foreign)
View(NHANESResearch)
head(NHANESResearch)
library(ggplot2)
library(epitools)
library(ggmosaic)
```

Descriptive statistics: You should generate descriptive statistics for each variable
included in your paper. The descriptive statistics that you use should be appropriate
for each kind of variable (e.g, nominal, ordinal, and continuous). You should also
illustrate the descriptive statistics for each variable via appropriate graphs.

```{r}

ggplot(NHANESResearch, aes(x=SMSpend)) + geom_histogram(fill = "turquoise") + labs(title = "Supermarket Spending", x="Supermarket Spending", y="count")
mean(NHANESResearch$SMSpend)
median(NHANESResearch$SMSpend)
sd(NHANESResearch$SMSpend)
range(NHANESResearch$SMSpend)

ggplot(NHANESResearch, aes(x=PovRatio)) + geom_histogram(fill = "turquoise") + labs(title = "Poverty Ratio", x="Poverty Ratio", y="count")
mean(NHANESResearch$SMSpend)
median(NHANESResearch$SMSpend)
sd(NHANESResearch$SMSpend)
range(NHANESResearch$SMSpend)

Race_table <- with(NHANESResearch, table(Race))
prop.table(Race_table)
Race_barchart <- as.data.frame(prop.table(Race_table)*100)
Race_barchart
ggplot(Race_barchart, aes(Race,Freq)) +geom_col(fill="turquoise")

Disability_table <- with(NHANESResearch, table(Disability))
prop.table(Disability_table)
Disability_barchart <- as.data.frame(prop.table(Disability_table)*100)
Disability_barchart
ggplot(Disability_barchart, aes(Disability,Freq)) +geom_col(fill="turquoise")

HHIncome_table <- with(NHANESResearch, table(HHIncome))
prop.table(HHIncome_table)
HHIncome_barchart <- as.data.frame(prop.table(HHIncome_table)*100)
HHIncome_barchart
ggplot(HHIncome_barchart, aes(HHIncome,Freq)) +geom_col(fill="turquoise")

FoodSec_table <- with(NHANESResearch, table(FoodSec))
prop.table(FoodSec_table)
FoodSec_barchart <- as.data.frame(prop.table(FoodSec_table)*100)
FoodSec_barchart
ggplot(FoodSec_barchart, aes(FoodSec,Freq)) +geom_col(fill="turquoise")

RaceD_table <- with(NHANESResearch, table(race_d))
prop.table(RaceD_table)
RaceD_barchart <- as.data.frame(prop.table(RaceD_table)*100)
RaceD_barchart
ggplot(RaceD_barchart, aes(race_d,Freq)) +geom_col(fill="turquoise")

```

Bivariate statistics: Your analysis must include a minimum of the following
statistical tests:
• One T-test
• One Chi-Square test
• One ANOVA or Kruskal-Wallis/Dunn test
• One correlation
• One linear regression model
• For each of these statistical tests, you should provide at least one table or graph
that illustrates these findings.
T-test
Compare the means of Poverty Ratio by White disabled and non- white disabled
```{r}
NHANESResearch%>%
  drop_na(Disability)%>%
  group_by(Disability)%>%
  summarize(mean(PovRatio), sd(PovRatio), count=n())
var.test(PovRatio~Disability, data=NHANESResearch, alternative="two.sided")
t.test(PovRatio~Disability, var.equal=FALSE, data=NHANESResearch)
ggplot(NHANESResearch, aes(x=Disability, y=PovRatio)) + geom_boxplot(fill="Red") + labs(x="Disability", y="Poverty Ratio", title="Disability by Poverty Ratio")
```
The p value is significant. 

Chi square
Compare the frequency of lower than pov ratio by white disabled and non- white disabled
```{r}
RD_FS_table <- xtabs(~ race_d + FoodSec, data=NHANESResearch)
RD_FS_table
prop.table(RD_FS_table)*100
chisq.test(RD_FS_table)
ggplot(NHANESResearch, aes(x=FoodSec, fill=race_d)) + geom_bar(position = "stack") + labs(title="Race with Disability by FoodSec", x="FoodSec", y="RaceD")
```
The p value is significant. 

Kruskal-Wallis/Dunn test
Supermarket spending by White disabled and non- white disabled
```{r}
#ggplot(NHANESdata,aes(x=BMIcat,y=DR1TKCAL))+geom_boxplot(fill="green")+labs(x="BMI category", y= "calories consumed per day", title = "Daily calories by BMIcat")
library(dunn.test)
dunn.test(x=NHANESResearch$SMSpend, g=NHANESResearch$race_d, kw=TRUE)
```
The P value is significant, but only for certain relationships. Neither disability or race are unilaterally factors in supermarket spending but there are significant differences.

Correlation
Disability by povratio for white and non- white
```{r}
#scatter plot code: 
ggplot(aes(x=PovRatio, y=SMSpend), data=NHANESResearch) + geom_point(color="red") + geom_smooth(method=lm, color="blue") + labs(x="PovRatio", y="Supermarket Spending") + theme_bw()

#correlation test code: 
cor.test(NHANESResearch$SMSpend, NHANESResearch$PovRatio, method="pearson")
```
The p value is signifcant and there is a weak positive relationship.

Linear regression model
```{r}
RaceD_PovRatio_fit <- lm(PovRatio~race_d, data=NHANESResearch)
summary(RaceD_PovRatio_fit)
ggplot(aes(x=race_d, y=PovRatio), data=NHANESResearch) + geom_point(color="red") + geom_smooth(method=lm, color="blue") + labs(x="Race with Disability", y="PovRatio") + theme_bw()
```
The p value is significant. There is a decreasing correlation between disability status within race and scale on the poverty ratio. There is a large positive relationship between asian disabled and poverty ratio than other races.

Odds Ratio: Your analysis should include at least one odds ratio that explains the
odds of developing some aspect of your dependent variable
Likelihood of being disabled by race
```{r}
#1. Create the odds table:
Disability_Race_table <- table(NHANESResearch$Race, NHANESResearch$Disability)
Disability_Race_table
#2. Run the oddsratio on the table you created:
oddsratio(Disability_Race_table)
#3. Graph the results in a mosaic table:
ggplot(data=NHANESResearch) + geom_mosaic(aes(x=product(Race), fill=Disability), na.rm=TRUE, show.legend=FALSE) + labs(title="Odds of Disability by Race", x="Disability", y="Race")
```
P value is significant only for Non- Hispanic White, Non-Hispanic Asian and Multiracial with an odds ratio of 0.54, 1.83, and 0.49 respectively.
